Eye in a Disk: eyeIntegration human pan-eye and body transcriptome database version 1.0

### 2019-05-07 manuscript.Rmd
adding sup figures, editing tables, small typo fixes

To better meet the needs of the eye research community we have re-written the bioinformatic pipeline that creates the eye and body RNA-seq dataset to allow for regular, versioned updates for eyeIntegration. We call this reproducible and versioned transcriptome dataset "Eye in a Disk" (EiaD). The pipeline automates the EiaD creation, ensures full reproducibility of the results, allow for external data comparison, provides consistent sample quality control, and improves efficiency for future sample updates. The 2019 EiaD dataset contains several new tissue types, full gene product quantification, along with over 100 new samples and improved sample labeling. The eyeIntegration web app has also been re-written to provide many new features, including versioned EiaD datasets, custom URL shortcut creation, new visualizations, improved data tables searching, easy download of core datasets, and local install of the entire app with three commands. Additionally, we are prototyping new tools to display single cell RNA-seq (scRNA-seq) data to provide researchers access to cell type specific information about gene expression across murine retinal development.

@@

## Quality control 
We first removed samples with a Salmon calculated mapping rate less than 40%. We further filtered samples by calculating the median gene expression for each sample and only kept samples that had a median count >2 across all genes. We removed lowly expressed genes by calculating the median expression across all samples for each gene and kept genes that had a median count >200 across all samples. To reduce the noise from experimental variability between each study, we normalized samples by sequence library size using the calcNormFactors function from the edgeR r-package, and then quantile smoothed expression data using the R package qsmooth at the tissue level [@hicksSmoothQuantileNormalization2018a; @robinsonEdgeRBioconductorPackage2010].

@@

@@

core_both %>% filter(grepl('Retina|Cornea|Lens|RPE', Tissue)) %>% unique() %>% 
  #mutate(Sub_Tissue = case_when(sample_accession == 'SRS1955479' ~ 'Retina - Adult Tissue', TRUE ~ Sub_Tissue)) %>%
  group_by(Tissue, Sub_Tissue, Version) %>% summarise(Count=n()) %>% 
  mutate(Dataset = Version) %>% 
  ggplot(aes(x=Sub_Tissue, y=Count, fill=Tissue, alpha = Dataset)) + geom_bar(stat = 'identity', position = 'dodge') + theme_minimal() + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.3)) + scale_colour_manual(values = c('grey','black')) + xlab('') +
  ylab('Count of Samples') + scale_alpha_manual(values = c(0.5,1)) +
  theme(text = element_text(size = 10)) +
  scale_fill_discrete_sequential(palette = 'viridis')  
```

@@

```{r workflowDoodle, echo = F, fig.width = 10, fig.height = 10, results='asis', fig.cap = workflowDoodle_cap}
workflowDoodle_cap <- fig_cap(name='workflowDoodle', caption = paste0('Raw RNA-seq data from the SRA is run through our pipeline to create the EiaD, which is used by eyeIntegration app to serve interactive gene expression visualizations across ', core_tight_2019 %>% filter(Kept == 'Kept') %>% pull(Sub_Tissue) %>% unique() %>% length(), ' tissues'))

@@

We built an automated pipeline for processing and analyzing all data for the web app using the program Snakemake, a python-based workflow management system that allows for efficient parallel execution of the analysis, facilitates reproduction by others, and simplifies long term maintenance of the EiaD data (`r fig_cap(name='workflowDoodle', display = 'cite')`, `r supFig_cap(name='snakemakeDAG', display = 'cite')`)[@kosterSnakemakeScalableBioinformatics2012]. To create a high quality final dataset across the `r core_tight_2019 %>% nrow()` initial samples (`r supTab_cap(name = 'metadataTable', display = 'cite')`) and 46,030,412,878 reads we developed a rigorous quality control procedure as part of our analysis, considering a sample’s read mapping rate and median count level as well as behavior relative to samples of the same sub-tissue type (see Methods).  To identify tube mislabeling or sample extraction issues, we used t-SNE to group samples by sub-tissue type to identify variability within samples of the same sub-tissue and ensure overall consistency in data processing (`r fig_cap(name = 'tSNE', display = 'cite')`).

@@

tsne_50_prep %>% mutate(Group = case_when(study_accession == 'SRP012682' ~ 'GTEx',
                                          TRUE ~ Tissue),
                        Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label),
                        Group = factor(Group, levels = c('Cornea', 'ESC', 'Lens', 'Retina', 'RPE','GTEx'))) %>% 
  ggplot(aes(x=X1,y=X2)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  geom_point(size=16, alpha=0.1, aes(colour=Group)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  xlab('t-SNE 1') + ylab('t-SNE 2') +
  geom_label_repel(size = 7, aes(label=Label), alpha=0.7, box.padding = unit(0.3, "lines")) + 
  theme_bw() +
  theme(text = element_text(size = 30)) +
  scale_color_discrete_sequential(palette = 'viridis') + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

@@

colnames(gtex) <- c('Group','Pre QC Count', 'Count', 'Tissue Types (Count)')
# Output <div>
cat("<div id=\"tbl:tableSampleCounts\">")
regulartable(table) %>% 
  width(width = c(1,1,1,2)) %>% 
  flextable::align_text_col(align = "center", header = TRUE) %>% 
  flextable::align(align='center') %>% 
  fontsize(size = 10) %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
# Output Caption (* * for italics)
cat(paste0("\n*",tableSampleCounts_cap,"*\n"))
# Output </div>
cat("</div>")
```

@@

## Custom gene and tissue expression boxplots
The 'Expression' tab of the webpage provides a wealth of information about both gene- and transcript-level expression for eye and body tissues, giving the user the ability to compare the expression of different genes within a single tissue, as well as the expression of genes across multiple tissues  (`r paste0(fig_cap(name = 'appScreenshots', display = 'cite'),'A')`). The user first selects either the 2017 or 2019 gene or transcript EiaD dataset, then genes (or transcripts), then tissues. A boxplot is then generated after hitting the "Re(Draw) Plot" button with overlaid individual data points. On mouse-over, the metadata for the individual sample is displayed. A tabular report is generated based on selected genes and tissues: a table with links to Ensembl, GeneCard, and OMIM for each gene for quick referencing, and a table containing expression levels for each selected gene in each selected tissue. The tables can be arranged or sorted to the user’s preference and can be easily downloaded for local use.

@@

@@

colors = viridis(n=6, direction = -1)
colors[5] <- "red"

# pca
make_pca_tissue <- function(tissue = 'Brain'){
  ggplot(dimRed_plot_prep_pca %>% 
           mutate(Origin = case_when(study_accession == 'SRP012682' ~ 'Adult Tissue',
                                     TRUE ~ as.character(Origin))) %>% 
           filter(!is.na(Origin)) %>% 
           mutate(Group = case_when(Tissue == !!(tissue) ~ Tissue,
                                    study_accession == 'SRP012682' ~ 'GTEx',
                                    TRUE ~ Tissue)) %>% 
           mutate(Group = factor(Group, levels= c('Cornea', 'ESC', 'Retina', 'RPE', !!(tissue), 'GTEx'))),
         aes(x=`Dimension 1`,y=`Dimension 2`)) +
    scale_shape_manual(values=c(0:20,35:50)) +
    #ggtitle(paste0('All Tissue PCA: Highlight ', tissue)) +
    geom_point(size=6, alpha=0.1, aes(colour=Group)) +
    geom_point(size=1, alpha=0.6, position='jitter', aes(shape=Origin)) +
    xlab('PCA 1') + ylab('PCA 2') +
    theme_bw() +
    # theme(text = element_text(family = 'Linux Libertine O', size = 10)) +
    scale_color_manual(values = colors) + 
    guides(colour = guide_legend(override.aes = list(alpha = 1)))
}
theme_set(theme_cowplot(font_size=10)) 
plot_grid(make_pca_tissue('Brain'), make_pca_tissue('Pituitary'),
          make_pca_tissue('Blood'), make_pca_tissue('Skin'),
          labels = c('A: Brain','B: Pituitary','C: Blood','D: Skin'),
          align = 'hv')
```

@@

```{r tableSampleCountsGTEx, echo = F, results='asis', warning = F, fig.cap = tableSampleCountsGTEx_cap, fig.pos = 'H'}
tableSampleCountsGTEx_cap <- supTab_cap(name='tableSampleCountsGTEx', caption = 'EiaD holds hundreds of GTEx tissues to provide a broad comparison set')
# Output <div>
cat("<div id=\"tbl:tableSampleCountsGTEx\">")
regulartable(gtex) %>% 
  width(width = c(0.5,1,0.5,3)) %>% 
  flextable::align_text_col(align = "center", header = TRUE) %>% 
  flextable::align(align='center') %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
#Output Caption (* * for italics)
cat(paste0("\n*",tableSampleCountsGTEx_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r tx_and_gene_types, echo = F, results='asis', warning = F, fig.cap =  tx_and_gene_types_cap, fig.pos = 'H'}
tx_and_gene_types_cap <- supTab_cap(name='tx_and_gene_types', caption = 'Dozens of different types of gene and transcript types quantified')
# Output <div>
cat("<div id=\"tbl:tx_and_gene_types\">")
table <- bind_rows(
  tx_anno %>% group_by(transcript_type) %>% summarise(Count=n()) %>% mutate(Type = 'Transcript') %>% rename(Class = transcript_type), 
  gene_anno %>% group_by(gene_type) %>% summarise(Count=n()) %>% mutate(Type = 'Gene') %>% rename(Class = gene_type))
regulartable(table %>% group_by(Type) %>% arrange(-Count)) %>% 
  width(width = c(2,1,2)) %>% 
  flextable::align_text_col(align = "center", header = TRUE) %>% 
  flextable::align(align='right') %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
#Output Caption (* * for italics)
cat(paste0("\n*",tx_and_gene_types_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r metadataTable, echo = F,  results='asis', fig.pos = 'H', fig.cap = metadata_cap}
metadata_cap <- supTab_cap(name='metadataTable', caption = 'Full metadata for 20 random eye samples. Full metadata available as supplementary file.')
set.seed(1534)
cat("<div id=\"tbl:metadataTable\">")
flextable(core_tight_2019 %>% filter(study_accession != 'SRP012682') %>% sample_n(20)) %>% 
  font(fontname = 'Linux Libertine O', part = 'all') %>% 
  fontsize(size = 6) %>% 
  width(width = c(0.5,0.5,0.5,3,1,0.5,0.5,0.5,0.5,0.5))
#Output Caption (* * for italics)
cat(paste0("\n*",metadata_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r GOterm_fetal_retina_organoid, echo = F, results='asis', warning = F, fig.cap = GOterm_cap, fig.pos = 'H'}
GOterm_cap <- supTab_cap(name='GOterm_fetal_retina_organoid', caption = 'Top GO terms enriched between fetal retina and organoid retina, after paring of redundant terms with REVIGO')
# Output <div>
cat("<div id=\"tbl:GOterm_fetal_retina_organoid\">")
organoid_fetal_GO %>% 
  dplyr::select(-ONTOLOGY, -BgRatio, -p.adjust, -Count, -Test, -Set) %>% 
  filter(ID %in% c('GO:0001503','GO:0031589','GO:0050878','GO:0061448','GO:0050678','GO:0030335','GO:0030198', 'GO:0007156', 'GO:0050769', 'GO:0030099')) %>% 
  mutate(geneID = gsub('<br>', ' ', geneID)) %>% 
  as_tibble() %>% 
  regulartable() %>% 
  flextable::fontsize(size = 8) %>% 
  width(width = c(1,1,1,1,1,2)) %>% 
  flextable::align_text_col(align = "center", header = TRUE) %>% 
  flextable::align(align='center') %>% 
  font(fontname = 'Linux Libertine O', part = 'all') 
#
#Output Caption (* * for italics)
cat(paste0("\n*",GOterm_cap,"*\n"))
# Output </div>
cat("</div>")
```
# Acknowledgements

>>>>

To better meet the needs of the eye research community we have re-written the bioinformatic pipeline that creates the eye and body RNA-seq dataset to allow for regular, versioned updates for eyeIntegration. We call this reproducible and versioned transcriptome dataset "Eye in a Disk" (EiaD). The pipeline automates the EiaD creation, ensures full reproducibility of the results, allow for external data comparison, provides consistent sample quality control, and improves efficiency for future sample updates. The 2019 EiaD dataset contains several new tissue types, full gene product quantification, along with over 100 new samples and improved sample labeling. The eyeIntegration web app has also been re-written to provide many new features, including versioned EiaD datasets, custom URL shortcut creation, new visualizations, improved data table searching, easy download of core datasets, and local install of the entire app with three commands. Additionally, we are prototyping new tools to display single cell RNA-seq (scRNA-seq) data to provide researchers access to cell type specific information about gene expression across murine retinal development.

@@

## Quality control 
We first removed samples with a Salmon calculated mapping rate less than 40%. This value was selected as being the far left tail of the distribution of mapping rates across samples (`r supFig_cap(name = 'mapping_rate_cap', display = 'cite')`). We further filtered samples by calculating the median gene expression for each sample and remove outlier samples (data not shown) that had a median count <2 across all genes. We removed lowly expressed genes by calculating the median expression across all samples for each gene and kept genes that had a median count >200 across all samples. To reduce the noise from experimental variability between each study, we normalized samples by sequence library size using the calcNormFactors function from the edgeR R package, and then quantile smoothed expression data using the R package qsmooth at the tissue level [@hicksSmoothQuantileNormalization2018a; @robinsonEdgeRBioconductorPackage2010].

@@

```{r, echo = F}
# set color levels
color <- c('#7fc97f','#beaed4','#ffff99','#fdc086','#386cb0','#f0027f', '#000000', '#000000', '#000000', '#000000')
names(color) <- c('Cornea', 'ESC', 'Lens', 'Retina', 'RPE', 'GTEx', 'Brain', 'Pituitary', 'Blood', 'Skin')
custom_fill <- scale_fill_manual(values = color)
custom_col <- scale_colour_manual(values = color)
```

@@

core_both %>% filter(grepl('Retina|Cornea|Lens|RPE', Tissue)) %>% unique() %>% 
  #mutate(Sub_Tissue = case_when(sample_accession == 'SRS1955479' ~ 'Retina - Adult Tissue', TRUE ~ Sub_Tissue)) %>%
  group_by(Tissue, Sub_Tissue, Version) %>% summarise(Count=n()) %>% 
  mutate(Dataset = Version) %>% 
  ggplot(aes(x=Sub_Tissue, y=Count, fill = Tissue, alpha = Dataset)) + geom_bar(stat = 'identity', position = 'dodge') + theme_minimal() + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.3))  + xlab('') +
  ylab('Count of Samples') + scale_alpha_manual(values = c(0.5,1)) +
  theme(text = element_text(size = 10)) +
  custom_fill
```

@@

```{r workflowDoodle, echo = F, fig.width = 10, fig.height = 10, results='asis', fig.cap = workflowDoodle_cap}
workflowDoodle_cap <- fig_cap(name='workflowDoodle', caption = paste0('Raw RNA-seq data from the SRA is run through our pipeline to create the EiaD, which is used by eyeIntegration app to serve interactive gene expression visualizations across ', core_tight_2019 %>% filter(Kept == 'Kept') %>% pull(Sub_Tissue) %>% unique() %>% length() - 1, ' tissues'))

@@

We built an automated pipeline for processing and analyzing all data for the web app using the program Snakemake, a python-based workflow management system that allows for efficient parallel execution of the analysis, facilitates reproduction by others, and simplifies long term maintenance of the EiaD data (`r fig_cap(name='workflowDoodle', display = 'cite')`, `r supFig_cap(name='snakemakeDAG', display = 'cite')`)[@kosterSnakemakeScalableBioinformatics2012]. To create a high quality final dataset across the `r core_tight_2019 %>% nrow()` initial samples (`r supTab_cap(name = 'metadataTable', display = 'cite')`) and 46,030,412,878 reads we developed a rigorous quality control procedure as part of our analysis, considering a sample’s read mapping rate and median count level as well as behavior relative to samples of the same sub-tissue type (see Methods).  To identify tube mislabeling or sample extraction issues, we used t-SNE to group samples by sub-tissue type to identify variability within samples of the same sub-tissue and ensure overall consistency in data processing (`r fig_cap(name = 'tSNE', display = 'cite')`). After these steps `r core_tight_2019 %>% filter(Kept != 'Kept') %>% filter(study_accession != 'SRP012682') %>% nrow()` eye samples and `r core_tight_2019 %>% filter(Kept != 'Kept') %>% filter(study_accession == 'SRP012682') %>% nrow()` GTEx samples were removed.

@@

tsne_50_prep %>% mutate(Group = case_when(study_accession == 'SRP012682' ~ 'GTEx',
                                          TRUE ~ Tissue),
                        Label = case_when(grepl('^0', Label) ~ '', TRUE ~ Label),
                        Group = factor(Group, levels = c('Cornea', 'ESC', 'Lens', 'Retina', 'RPE','GTEx'))) %>% 
  ggplot(aes(x=X1,y=X2)) +
  scale_shape_manual(values=c(0:20,35:50)) +
  geom_point(size=16, alpha=0.1, aes(colour=Group)) +
  geom_point(size=6, alpha=0.6, position='jitter', aes(shape=Origin)) +
  xlab('t-SNE 1') + ylab('t-SNE 2') +
  geom_label_repel(size = 7, aes(label=Label), alpha=0.7, box.padding = unit(0.3, "lines")) + 
  theme_bw() +
  theme(text = element_text(size = 30)) +
  custom_col + 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

@@

colnames(gtex) <- c('Group','Pre QC Count', 'Count', 'Tissue Types (Count)')
# Output <div>
cat("<div id=\"tbl:tableSampleCounts\">")
regulartable(table) %>% 
  width(width = c(1,1,1,2)) %>% 
  flextable::align(align = "center",part = "header") %>% 
  flextable::align(align='center') %>% 
  fontsize(size = 10) %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
# Output Caption (* * for italics)
cat(paste0("\n*",tableSampleCounts_cap,"*\n"))
# Output </div>
cat("</div>")
```

@@

## Custom gene and tissue expression boxplots
The 'Expression' tab of the webpage provides a wealth of information about both gene- and transcript-level expression for eye and body tissues, giving the user the ability to compare the expression of different genes within a single tissue, as well as the expression of genes across multiple tissues  (`r paste0(fig_cap(name = 'appScreenshots', display = 'cite'),'A')`). The user first selects either the 2017 or 2019 gene or transcript EiaD dataset, then Hugo Gene Nomeclature Committee (HGNC) genes names (or ENSEMBL transcripts), then tissues. A boxplot is then generated after hitting the "Re(Draw) Plot" button with overlaid individual data points. On mouse-over, the metadata for the individual sample is displayed. A tabular report is generated based on selected genes and tissues: a table with links to Ensembl, GeneCard, and OMIM for each gene for quick referencing, and a table containing expression levels for each selected gene in each selected tissue. The tables can be arranged or sorted to the user’s preference and can be easily downloaded for local use.

@@

```{r mapping_rate, echo = F, message = F,  warning=F, fig.height = 4, fig.width = 10, fig.pos='H', fig.cap = mapping_rate_cap}
mapping_rate <- read_delim('data/mapping_rate.txt', delim = ' ', col_names = F) %>% 
  mutate(X2 = str_sub(X2, 1, 4) %>% as.numeric()) %>%  
  rename(sample_accession = X1, `Mapping Rate` = X2)
mapping_stats <- core_tight_2019 %>% left_join(mapping_rate) %>% pull(`Mapping Rate`) %>% summary() %>% broom::tidy()
mapping_rate_cap <- supFig_cap(name='mapping_rate_cap', caption = paste0('Salmon mapping rate for each sample, grouped by tissue type. 1st quartile mapping rate is ', mapping_stats$q1, '%, median is ', mapping_stats$median, '%, mean is ', mapping_stats$mean, '%, and third quartile is ', mapping_stats$q3, '%.'))
core_tight_2019 %>% 
  left_join(mapping_rate) %>% 
  ggplot(aes(x=Tissue, y = `Mapping Rate`)) + 
  ggbeeswarm::geom_quasirandom() + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_cartesian(ylim=c(0,100))

@@

# colors = viridis(n=6, direction = -1)
# colors[5] <- "red"

# pca
make_pca_tissue <- function(tissue = 'Brain'){
  ggplot(dimRed_plot_prep_pca %>% 
           mutate(Origin = case_when(study_accession == 'SRP012682' ~ 'Adult Tissue',
                                     TRUE ~ as.character(Origin))) %>% 
           filter(!is.na(Origin)) %>% 
           mutate(Group = case_when(Tissue == !!(tissue) ~ Tissue,
                                    study_accession == 'SRP012682' ~ 'GTEx',
                                    TRUE ~ Tissue)) %>% 
           mutate(Group = factor(Group, levels= c('Cornea', 'ESC', 'Retina', 'RPE', !!(tissue), 'GTEx'))),
         aes(x=`Dimension 1`,y=`Dimension 2`)) +
    scale_shape_manual(values=c(0:20,35:50)) +
    #ggtitle(paste0('All Tissue PCA: Highlight ', tissue)) +
    geom_point(size=6, alpha=0.1, aes(colour=Group)) +
    geom_point(size=1, alpha=0.8, position='jitter', aes(shape=Origin)) +
    xlab('PCA 1') + ylab('PCA 2') +
    theme_bw() +
    # theme(text = element_text(family = 'Linux Libertine O', size = 10)) +
    theme_minimal() +
    custom_col + 
    guides(colour = guide_legend(override.aes = list(alpha = 1)))
}
theme_set(theme_cowplot(font_size=10)) 
plot_grid(make_pca_tissue('Brain'), make_pca_tissue('Pituitary'),
          make_pca_tissue('Blood'), make_pca_tissue('Skin'),
          labels = c('A: Brain','B: Pituitary','C: Blood','D: Skin'),
          align = 'hv')
```

@@

```{r tableSampleCountsGTEx, echo = F, results='asis', warning = F, fig.cap = tableSampleCountsGTEx_cap, fig.pos = 'H'}
tableSampleCountsGTEx_cap <- supTab_cap(name='tableSampleCountsGTEx', caption = 'EiaD holds hundreds of GTEx tissues to provide a broad comparison set')
# Output <div>
cat("<div id=\"tbl:tableSampleCountsGTEx\">")
regulartable(gtex) %>% 
  width(width = c(0.5,1,0.5,3)) %>% 
  flextable::align(align = "center", part = 'header') %>% 
  flextable::align(align='center') %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
#Output Caption (* * for italics)
cat(paste0("\n*",tableSampleCountsGTEx_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r tx_and_gene_types, echo = F, results='asis', warning = F, fig.cap =  tx_and_gene_types_cap, fig.pos = 'H'}
tx_and_gene_types_cap <- supTab_cap(name='tx_and_gene_types', caption = 'Dozens of different types of gene and transcript types quantified')
# Output <div>
info <- readxl::read_xlsx('data/gencode_type_description.xlsx') %>% fill('Definition') %>% 
  mutate(Biotype = case_when(Biotype == 'antisense/antisense_RNA' ~ 'antisense',
                             TRUE ~ Biotype))
cat("<div id=\"tbl:tx_and_gene_types\">")
table <- bind_rows(
  tx_anno %>% group_by(transcript_type) %>% summarise(Count=n()) %>% mutate(Type = 'Transcript') %>% rename(Class = transcript_type), 
  gene_anno %>% group_by(gene_type) %>% summarise(Count=n()) %>% mutate(Type = 'Gene') %>% rename(Class = gene_type))
table <- table %>% 
  filter(!is.na(Class)) %>% 
  rename(Biotype = Class) %>% 
  spread(Type, Count) %>% 
  rename(`Gene Count` = Gene, `Transcript Count` = Transcript) %>% 
  left_join(info)
regulartable(table%>% arrange(Definition)) %>% 
  width(width = c(2,1,1,4)) %>% 
 flextable::align(align = "center", part = 'header') %>% 
  flextable::align(align='right') %>% 
  font(fontname = 'Linux Libertine O', part = 'all')
#Output Caption (* * for italics)
cat(paste0("\n*",tx_and_gene_types_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r metadataTable, echo = F,  results='asis', fig.pos = 'H', fig.cap = metadata_cap}
metadata_cap <- supTab_cap(name='metadataTable', caption = 'Full metadata for 20 random eye samples. Full metadata available as supplementary file.')
set.seed(1534)
cat("<div id=\"tbl:metadataTable\">")
flextable(core_tight_2019 %>% filter(study_accession != 'SRP012682') %>% sample_n(20) %>% left_join(mapping_rate) %>% rename(Mapping_Rate = `Mapping Rate`)) %>% 
  font(fontname = 'Linux Libertine O', part = 'all') %>% 
  fontsize(size = 6) %>% 
  width(width = c(0.4,0.4,0.4,2.8,1,0.4,0.4,0.4,0.4,0.4,0.4))
#Output Caption (* * for italics)
cat(paste0("\n*",metadata_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r GOterm_fetal_retina_organoid, echo = F, results='asis', warning = F, fig.cap = GOterm_cap, fig.pos = 'H'}
GOterm_cap <- supTab_cap(name='GOterm_fetal_retina_organoid', caption = 'Top GO terms enriched between fetal retina and organoid retina, after paring of redundant terms with REVIGO')
# Output <div>
cat("<div id=\"tbl:GOterm_fetal_retina_organoid\">")
organoid_fetal_GO %>% 
  dplyr::select(-ONTOLOGY, -BgRatio, -p.adjust, -Count, -Test, -Set) %>% 
  filter(ID %in% c('GO:0001503','GO:0031589','GO:0050878','GO:0061448','GO:0050678','GO:0030335','GO:0030198', 'GO:0007156', 'GO:0050769', 'GO:0030099')) %>% 
  mutate(geneID = gsub('<br>', ' ', geneID)) %>% 
  as_tibble() %>% 
  regulartable() %>% 
  flextable::fontsize(size = 8) %>% 
  width(width = c(1,1,1,1,1,2)) %>% 
  flextable::align(align = "center", part = 'header') %>% 
  flextable::align(align='center') %>% 
  font(fontname = 'Linux Libertine O', part = 'all') 
#
#Output Caption (* * for italics)
cat(paste0("\n*",GOterm_cap,"*\n"))
# Output </div>
cat("</div>")
```
# Acknowledgements
